---
title: 'MiST: *Mi*xed effects *S*core *T*est'
author: "Jianping Sun, Yingye Zheng and Li Hsu"
email: "mickael.canouil@cnrs.fr"
date: '`r format(Sys.time(), "%d %B %Y")`'
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
    fig_width: 6.3
    fig_height: 4.7
vignette: >
  %\VignetteIndexEntry{MiST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  collapse = TRUE,
  include = TRUE,
  echo = TRUE,
  warning = TRUE,
  message = TRUE,
  error = TRUE,
  fig.align = 'center',
  fig.pos = '!h',
  cache = FALSE
)
```

Test for association between a set of SNPS/genes and continuous or binary outcomes by including variant characteristic information and using (weighted) score statistics.

**Note:**

* From: https://cran.r-project.org/package=MiST (v1.0)
* Reference: https://doi.org/10.1002/gepi.21717

## Installation

```{r, eval = FALSE}
# Install MiST from CRAN:
install.packages("MiST")

# Or the the development version from GitHub:
# install.packages("remotes")
remotes::install_github("mcanouil/MiST")
```

## MiST in Action

```{r}
library(MiST)
data(mist_data)
attach(mist_data)
```

### Continuous Outcome

#### With Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_tau"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous"
)
```

#### With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_tau"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "continuous"
)
```

#### With "Average" Effect (π)

```{r}
mist(
  y = phenotypes[, "y_pi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous"
)
```

#### With "Average" Effect (π) and Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_taupi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous"
)
```

#### With "Average" Effect (π) and With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_taupi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "continuous"
)
```

### Continuous Outcome with Weights

#### With Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_tau"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous", 
  weight.beta = c(1, 25), 
  maf = variants_info[, "maf"]
)
```

#### With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_tau"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "continuous", 
  weight.beta = c(1, 25), 
  maf = variants_info[get_same_effect, "maf"]
)
```

#### With "Average" Effect (π)

```{r}
mist(
  y = phenotypes[, "y_pi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous", 
  weight.beta = c(1, 25), 
  maf = variants_info[, "maf"]
)
```

#### With "Average" Effect (π) and Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_taupi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "continuous", 
  weight.beta = c(1, 25), 
  maf = variants_info[, "maf"]
)
```

#### With "Average" Effect (π) and With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_taupi"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "continuous", 
  weight.beta = c(1, 25), 
  maf = variants_info[get_same_effect, "maf"]
)
```


### Binary Outcome

#### With Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_binary"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "binary"
)
```

#### With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_binary"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "binary"
)
```

### Binary Outcome with Weights

#### With Heterogeneity (τ)

```{r}
mist(
  y = phenotypes[, "y_binary"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes,
  Z = variants_info[, 1, drop = FALSE],
  method = "liu",
  model = "binary", 
  weight.beta = c(1, 25), 
  maf = variants_info[, "maf"]
)
```

#### With Heterogeneity (τ) "Removed"

```{r}
variants_info[, "effect"] # simulated effect
get_same_effect <- names(which(variants_info[, "effect"] > 0))
mist(
  y = phenotypes[, "y_binary"],
  X = phenotypes[, paste0("x_cov", 0:2)],
  G = genotypes[, get_same_effect],
  Z = variants_info[get_same_effect, 1, drop = FALSE],
  method = "liu",
  model = "binary", 
  weight.beta = c(1, 25), 
  maf = variants_info[get_same_effect, "maf"]
)
```
